---
title: "OLWTEF"
author: "JJayes"
date: "05/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)

rm(list = ls())

```


```{r cars}
library(pacman)
p_load(tidyverse, ggrepel)

theme_set(theme_classic())


# key_crop_yields <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/key_crop_yields.csv')
# 
# # clean names
# key_crop_yields <- key_crop_yields %>% 
#   janitor::clean_names() %>% 
#   rename_all(str_remove, "_tonnes.*")
# 
# # data processing
# yields_tidy <- key_crop_yields  %>% 
#   pivot_longer(wheat:bananas, names_to = "crop", values_to = "yield") %>% 
#   filter(!is.na(yield))
# 
# write_rds(yields_tidy, "yields_tidy.rds")

```

## How do we thrive?
### Share of population living in extreme poverty by region



```{r}
df <- read.csv("data/WDIData.csv")

df <- df %>% 
  janitor::clean_names()

df_pov <- df %>% 
  filter(indicator_code %in% c("SI.POV.DDAY", "SI.POV.MDIM", "SI.POV.LMIC.GP", "SI.POV.GINI")) %>% as_tibble()

df_pov <- df_pov %>% 
  pivot_longer(x1960:x2019, names_to = "year", values_to = "stat") %>% 
  select(-x) %>% 
  mutate(year = parse_number(year)) %>% 
  rename(country_name = i_country_name)

write_rds(df_pov, "data/df_pov.rds")


```

```{r}
# df_pov %>% 
#   filter(country_code %in% c("DOM", "DZA", "EAP")) %>% 
#   ggplot(aes(year, stat, lty = country_name)) +
#   geom_line() +
#   facet_wrap(~ indicator_name) + 
#   scale_color_brewer(type = "qual", name = "Country") +
#   theme(legend.position = "bottom")
# 
# g <- df_pov %>% 
#   filter(!is.na(stat)) %>% 
#   ggplot(aes(year, stat, group = country_code)) +
#   geom_line() +
#   facet_wrap(~ indicator_name)
# 
# # trying gghighlight
# df_pov %>% 
#   filter(!is.na(stat)) %>% 
#   ggplot() +
#   geom_line(aes(year, stat, colour = country_code)) +
#   gghighlight(country_code == "DOM") +
#   facet_wrap(~ indicator_name)

# data cleaning and processing for poverty data
df_pov <- readRDS("data/df_pov.rds")

df_pov <- df_pov %>% 
  filter(!is.na(stat))

df_pov_graph <- df_pov %>%
  filter(indicator_code == "SI.POV.DDAY",
         year <= 2016,
         country_name %in% c("South Africa", "Sub-Saharan Africa", "South Asia", "World", "Middle East & North Africa", "East Asia % Pacific", 
                                  #"Latin America & Caribbean", 
                                  "Europe & Central Asia")) %>% 
  mutate(country_name = fct_reorder(country_name, -stat))
```

Share of population living in extreme poverty by region: Thin dotted lines and legend

```{r}

df_pov_graph %>% 
  ggplot() +
  geom_line(aes(year, stat, lty = country_name)) +
  # geom_text_repel(data = df_pov_graph %>%
  #                   group_by(country_name) %>% 
  #                   filter(year == max(year)), 
  #                 aes(year, stat, label = country_name),
  #                 hjust = -1, vjust = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in extreme poverty by region",
       subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries \n It is measured in international-$ in 2011 prices",
       caption = "Source: World Bank World Development Indicators", linetype = "Region")

```

Share of population living in extreme poverty by region: Thick dotted lines and grey shading - labels in pane and no legend

```{r}

df_pov_graph %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name), cex = 1) +
  geom_text_repel(data = df_pov_graph %>% 
                    group_by(country_name) %>% 
                    filter(year == max(year)), 
                  aes(year, stat, label = country_name)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  scale_colour_grey(start = 0.8, end = 0.2) +
  expand_limits(x = 2020) +
  # scale_colour_brewer(type = "seq", palette = "Greys",  name = "Region") +
  theme(legend.position = "none") +
  labs(y = "", 
       x = "",
       title = "Share of population living in extreme poverty by region",
       subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries \n It is measured in international-$ in 2011 prices",
       caption = "Source: World Bank World Development Indicators")

```

Share of population living in extreme poverty by region: Shaded thick dotted line and a legend

```{r}

df_pov_graph %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name, lty = country_name), cex = 2) +
  # geom_text_repel(data = df_pov_graph %>%
  #                   group_by(country_name) %>% 
  #                   filter(year == max(year)), 
  #                 aes(year, stat, label = country_name),
  #                 hjust = -1, vjust = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  # scale_colour_brewer(type = "seq", palette = "Greys",  name = "Region") +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in extreme poverty by region",
       subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries \n It is measured in international-$ in 2011 prices",
       caption = "Source: World Bank World Development Indicators",
       colour = "Region", linetype = "Region")

```

Share of population living in extreme poverty by region: Thinner dotted line and a legend

```{r}
df_pov_graph %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name, lty = country_name), cex = .8) +
  # geom_text_repel(data = df_pov_graph %>%
  #                   group_by(country_name) %>% 
  #                   filter(year == max(year)), 
  #                 aes(year, stat, label = country_name),
  #                 hjust = -1, vjust = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  # scale_colour_brewer(type = "seq", palette = "Greys",  name = "Region") +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in extreme poverty by region",
       subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries \n It is measured in international-$ in 2011 prices",
       caption = "Source: World Bank World Development Indicators",
       colour = "Region", linetype = "Region")

```


..

All of the following plots can be done with any selection of these styles, as you like the look most.

..

### Share of population living in absolute poverty

Share of population living in absolute poverty: Thick, dotted grey lines

```{r}
df_pov_world <- df_pov_graph %>% 
  filter(country_name == "World") %>% 
  select(label = country_name,
         year,
         stat) %>% 
  mutate(label = fct_recode(label, "Living in poverty (World Bank Data) = on less than 1.90$ per day" = "World"))

library(readxl)

share_of_pop_in_absolute_poverty_bourguinon <- read_excel("data/share of pop in absolute poverty_bourguinon.xlsx")

df_pov_comp <- rbind(df_pov_world, share_of_pop_in_absolute_poverty_bourguinon)

df_pov_comp %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = label, lty = label), cex = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  expand_limits(y = 0) +
  # scale_colour_brewer(type = "seq", palette = "Greys") +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in absolute poverty, 1820-2015",
       # subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries - it is measured in international-$ in 2011 prices",
       lty = " ",
       colour = " ") +
  guides(colour = guide_legend(nrow=3,byrow=TRUE))

```

Share of population living in absolute poverty: Thin dotted lines

```{r}
df_pov_comp %>% 
  ggplot() +
  geom_line(aes(year, stat, lty = label), cex = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  expand_limits(y = 0) +
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in absolute poverty, 1820-2015",
       # subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries - it is measured in international-$ in 2011 prices",
       lty = " ") +
  guides(lty = guide_legend(nrow=3,byrow=TRUE))

```


Share of population living in absolute poverty: shaded lines

```{r}
df_pov_comp %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = label), cex = 1) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) + 
  expand_limits(y = 0) +
  scale_colour_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "bottom") +
  labs(y = "", 
       x = "",
       title = "Share of population living in absolute poverty, 1820-2015",
       # subtitle = "GDP per capita adjusted for price changes over time (inflation) and price differences between countries - it is measured in international-$ in 2011 prices",
       colour = " ") +
  guides(colour = guide_legend(nrow=3,byrow=TRUE))

```



```{r}

df_gdp <- df %>% 
  filter(indicator_code %in% c("NY.GDP.PCAP.KD", "NY.GDP.PCAP.KN", "NY.GDP.PCAP.CN", "NY.GDP.PCAP.CD")) %>% as_tibble()

df_gdp <- df_gdp %>% 
  pivot_longer(x1960:x2019, names_to = "year", values_to = "stat") %>% 
  select(-x) %>% 
  mutate(year = parse_number(year))

# write data
write_rds(df_gdp, "data/df_gdp.rds")
# 
# df_gdp %>%
#   filter(country_code %in% c("DOM", "DZA", "EAP")) %>%
#   ggplot(aes(year, stat, colour = i_country_name)) +
#   geom_line() +
#   facet_wrap(~ indicator_name, scales = "free_y") +
#   scale_color_brewer(type = "qual", name = "Country") +
#   theme(legend.position = "bottom")

# # gghighlight again
# df_gdp %>%
#   filter(indicator_code == "NY.GDP.PCAP.KD") %>%
#   ggplot() +
#   geom_line(aes(year, stat, colour = i_country_name)) +
#   gghighlight(i_country_name %in% c("South Africa", "China", "Brazil", "United States")) +
#   scale_y_continuous(labels = scales::comma_format()) +
#   labs(y = "GDP per capita (constant 2010 US$)",
#        x = "Year",
#        title = "GDP per capita growth since 1960")
# 
# p_load(devtools)
# devtools::install_github("yutannihilation/gghighlight")
library(gghighlight)


```

### GDP per capita growth

(Choice of countries here can be changed)

```{r}
df_mad <- read.csv("data/maddison_project_database.csv")

df_mad <- df_mad %>% 
  janitor::clean_names() %>% 
  as_tibble()

df_mad <- df_mad %>%
  pivot_longer(cgdppc:pop, names_to = "indicator_name", values_to = "stat") %>% 
  select(- i_cig, - i_bm)

df_mad <- df_mad %>%
  mutate(indicator_code = indicator_name,
         indicator_name = recode(indicator_code, pop = "Population", cgdppc ="Real GDP per capita in 2011US$ (for income comparisons)", rgdpnapc = "Real GDP per capita in 2011US$ (for growth comparisons)")) %>% 
  rename(country_code = i_countrycode,
         country_name = country)

write_rds(df_mad, "data/maddison_project_database.rds")
```

GDP per capita growth since 1700: Thin dotted lines, labels in pane and a legend.

```{r}
df_mad_plot <- df_mad %>%
  filter(country_name %in% c("South Africa", "China", "Botswana", "United Kingdom", "Zambia"),
         indicator_code == "rgdpnapc",
         year > 1700) 

df_mad_plot %>% 
  ggplot() +
  geom_line(aes(year, stat, lty = country_name)) +
  geom_text_repel(data = df_mad %>%
              filter(country_name %in% c("South Africa", "China", "Botswana", "United Kingdom", "Zambia"),
              indicator_code == "rgdpnapc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.4)) +
  scale_y_continuous(labels = scales::comma_format()) + 
  expand_limits(x = 2050) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       title = "GDP per capita growth since 1700",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "bottom",
        legend.title = element_blank())

```

GDP per capita growth since 1700: Grey shading and legend

```{r}
df_mad_plot %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name), cex = 1) +
  geom_text_repel(data = df_mad %>%
              filter(country_name %in% c("South Africa", "China", "Botswana", "United Kingdom", "Zambia"),
              indicator_code == "rgdpnapc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.4)) +
  # scale_color_brewer(palette = "Greys") +
  scale_colour_grey(start = 0.8, end = 0.2) +
  scale_y_continuous(labels = scales::comma_format()) + 
  expand_limits(x = 2050) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       title = "GDP per capita growth since 1700",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "bottom",
        legend.title = element_blank())

# with ggghighlight
# df_mad %>%
#   filter(indicator_code == "rgdpnapc",
#          stat < 50000,
#          year > 1000) %>% 
#   ggplot() +
#   geom_line(aes(year, stat, colour = country_name)) +
#   gghighlight(country_name %in% c("South Africa", "China", "Botswana", "United Kingdom", "Zambia")) + 
#   scale_y_continuous(labels = scales::comma_format()) + 
#   scale_colour_brewer(type = "qual", palette = "Dark2",  name = "Country") +
#   labs(y = "GDP per capita (constant 2010 US$)", 
#        x = "Year",
#        title = "GDP per capita growth since 1960",
#        caption = "Source: World Bank World Development Indicators")

```

GDP per capita growth since 1700: Grey shading and dotted lines - no legend

```{r}

df_mad_plot %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name, lty = country_name), cex = 1) +
  geom_text_repel(data = df_mad %>%
              filter(country_name %in% c("South Africa", "China", "Botswana", "United Kingdom", "Zambia"),
              indicator_code == "rgdpnapc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.4)) +
  scale_color_grey(start = 0.8, end = 0.2) +
  scale_y_continuous(labels = scales::comma_format()) + 
  expand_limits(x = 2050) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       title = "GDP per capita growth since 1700",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "none",
        legend.title = element_blank())

```



### Ashraf and Galor

Observed genetic heterogeneity and log population density

```{r}
library(readxl)
ashraf_and_galor <- read_excel("data/ashraf and galor.xlsx")

ashraf_and_galor %>% 
  ggplot(aes(x, y)) +
  geom_point(aes(x, y, shape = continent), cex = 2) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_text_repel(aes(x, y, label = country_name)) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), cex = 1, colour = "black") +
  scale_shape_manual(values = seq(15,19, by = 1)) +
  labs(x = "(Observed) Genetic homogeneity",
       y = "log population density in 1500 CE",
       caption = "Source: Ashraf and Galor (2013)") +
  theme_classic() +
  theme(legend.title = element_blank())  +
  theme(legend.position = "bottom")

```



### Timeline of Jerusalem

Timeline of Jerusalem: Dotted lines - labels on y-axis

(Can change source here if you would prefer not to cite Wikipedia)

```{r}
timeline_of_jerusalem <- read_excel("data/timeline of jerusalem.xlsx")

timeline_of_jerusalem <- timeline_of_jerusalem %>% 
  rowwise() %>% 
  mutate(mymean = mean(c(date_start,date_end) )) %>% 
  arrange(desc(mymean)) %>% 
  mutate(period = factor(period))

timeline_of_jerusalem %>% 
  ggplot() +
  geom_segment(aes(y =period, yend = period, x = date_start, xend = date_end, lty = religion)) +
  # geom_point(aes(x = date_start, y = period, shape = religion)) + 
  theme(legend.title = element_blank()) +
  theme(legend.position = "bottom") +
  labs(x = "Date", 
       y = "",
       caption = "Source: Wikipedia")

```

Timeline of Jerusalem: Shades of grey - labels on y-axis

```{r}

timeline_of_jerusalem %>% 
  ggplot() +
  geom_segment(aes(y =period, yend = period, x = date_start, xend = date_end, colour = religion), cex = 3) +
  # geom_point(aes(x = date_start, y = period, shape = religion)) + 
  theme(legend.title = element_blank()) +
  scale_color_grey(start = 0.8, end = 0.2) +
  theme(legend.position = "bottom") +
  labs(x = "Date", 
       y = "",
       caption = "Source: Wikipedia")

```

Timeline of Jerusalem: Alternate dotted lines - labels beside lines

```{r}
timeline_of_jerusalem %>% 
  ggplot() +
  geom_segment(aes(y =period, yend = period, x = date_start, xend = date_end, lty = religion), cex = 1) +
  geom_point(aes(x = date_start, y = period, shape = religion)) + 
  theme_classic() +
  geom_text(aes(x = date_end, y = period, label = period, hjust = -.1), cex = 3) +
  scale_shape_manual(values = seq(15,19, by = 1)) +
  scale_x_continuous(breaks = seq(-2000, 2000, by = 500)) +
  theme(legend.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "bottom") +
  labs(x = "", 
       y = "",
       caption = "Source: Wikipedia")
```

Timeline of Jerusalem:  Alternate shades of grey - labels beside lines.
Personally I like this one the most.

```{r}
timeline_of_jerusalem %>% 
  ggplot() +
  geom_segment(aes(y =period, yend = period, x = date_start, xend = date_end, colour = religion), cex = 3) +
  # geom_point(aes(x = date_start, y = period, shape = religion)) + 
  theme_classic() +
  geom_text(aes(x = date_end, y = period, label = period, hjust = -.1), cex = 3) +
  scale_color_grey(start = 0.8, end = .2) +
  scale_x_continuous(breaks = seq(-2000, 2000, by = 500)) +
  expand_limits(x = 2500) +
  # scale_shape_manual(values = seq(15,19, by = 1)) +
  theme(legend.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "bottom") +
  labs(x = "", 
       y = "",
       caption = "Source: Wikipedia")

# must fix y axis ticks

# make grey lines thicker and diverge by 5 colours
  
```

### Slave routes map

```{r}

library(tidygeocoder)

cities <- c("dakar, senegal",
             "freetown, sierra leone",
            "monrovia, liberia",
            "abidjan, cote d'ivoire",
            "lagos, nigeria",
            "douala, cameroon",
            "luanda, angola",
            "beira, mozambique") %>% as_tibble()

# cities_coords <- cities %>% geocode(address = value, method = "osm", lat = lat1, long = long1)
# 
# write_rds(cities_coords, "data/cities_coords.rds")

cities_coords <- read_rds("data/cities_coords.rds")

cities_coords <- cities_coords %>% 
  mutate(value = str_to_title(value))

senders <- c("rio de janeiro, brazil",
             "san juan, puerto rico") %>% as_tibble()

# senders_coords <- senders %>% geocode(address = value, method = "osm", lat = lat1, long = long1)
# 
# write_rds(senders_coords, "data/senders_coords.rds")

senders_coords <- read_rds("data/senders_coords.rds")

senders_coords <- senders_coords %>% 
  mutate(value = str_to_title(value))

carib <- senders_coords %>% filter(value == "San Juan, Puerto Rico")

brazil <- senders_coords %>% filter(value == "Rio De Janeiro, Brazil")

from_carib <- cities_coords %>% 
  filter(!value %in% c("beira, mozambique"))

from_brazil <- cities_coords %>% 
  filter(value %in% c("lagos, nigeria", 
                      "luanda, angola",
                      "beira, mozambique"))

carib_route <- from_carib %>% 
  mutate(lat2 = carib$lat1,
         long2 = carib$long1) %>% 
  select(-value)

brazil_route <- from_brazil %>% 
  mutate(lat2 = brazil$lat1,
         long2 = brazil$long1) %>% 
  select(-value)

```

```{r}
library(maps)

# No margin
par(mar=c(0,0,0,0))

p_load(geosphere)

coords <- rbind(cities_coords, senders_coords)

# Show the cities on the map

```

```{r}
# plot_my_connection = function( dep_lon, dep_lat, arr_lon, arr_lat, ...){
#     inter <- gcIntermediate(c(dep_lon, dep_lat), c(arr_lon, arr_lat), n=50, addStartEnd=TRUE, breakAtDateLine=F)             
#     inter=data.frame(inter)
#     diff_of_lon=abs(dep_lon) + abs(arr_lon)
#     if(diff_of_lon > 180){
#         lines(subset(inter, lon>=0), ...)
#         lines(subset(inter, lon<0), ...)
#     }else{
#         lines(inter, ...)
#         }
# }
# 
# map('world',
#     col="#f2f2f2", fill=TRUE, bg="white", lwd=0.05,
#     mar=rep(0,4),border=0, ylim=c(-60,40), xlim = c(-80,60)
# )
# 
# for(i in 1:nrow(carib_route)){
#     plot_my_connection(carib_route$long1[i], carib_route$lat1[i], carib_route$long2[i], carib_route$lat2[i], col="black", lwd=1)
# }
# 
# for(i in 1:nrow(brazil_route)){
#     plot_my_connection(brazil_route$long1[i], brazil_route$lat1[i], brazil_route$long2[i], brazil_route$lat2[i], col="black", lwd=1, lty = 2)
#     }
#  
# 
# 
# # add points and names of cities
# points(x=coords$long1, y=coords$lat1, col="black", cex=2, pch=18)
# 
# text(cities_coords$value, x=cities_coords$long1, y=cities_coords$lat1,  col="black", cex=.6, pos=4)
# 
# text(senders_coords$value, x=senders_coords$long1, y=senders_coords$lat1,  col="black", cex=.6, pos=1)
```

Slave routes map: No labels

```{r}

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = world) +
  geom_sf(colour = "white") +
  geom_point(data = coords, aes(long1, lat1), shape = 18, cex = 2) +
  coord_sf(ylim=c(-60,40), xlim = c(-80,60)) +
  geom_curve(data = carib_route,
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             curvature = -0.3,
             # arrow = arrow(length = unit(0.02, "npc"), type = "closed"), 
             lty = 2) +
  geom_curve(data = brazil_route, 
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             # arrow = arrow(length = unit(0.02, "npc"), type = "closed"), ) +
             lty = 1, curvature = 0.3) + 
  labs(x = "Longitude",
       y = "Latitude") +
  theme_void()

```

Slave routes map: Labels

```{r}

wa_labels <- c("Senegambia",
  "Sierra Leone",
  "Windward Coast",
  "Gold Coast",
  "Bight of Benin",
  "Bight of Biafra",
  "West Central Africa",
  "Southeast Africa",
  "Brazil",
  "Caribbean")

from_carib_labels <- cbind(coords, wa_labels)

ggplot(data = world) +
  geom_sf(colour = "white") +
  geom_point(data = coords, aes(long1, lat1), shape = 18, cex = 2) +
  coord_sf(ylim=c(-60,40), xlim = c(-80,60)) +
  geom_text_repel(data = from_carib_labels, aes(long1, lat1, label = wa_labels), hjust = -.1, vjust = -.3) +
  geom_curve(data = carib_route,
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             curvature = 0.3,
             # arrow = arrow(length = unit(0.02, "npc"), type = "closed"), 
             lty = 2) +
  geom_curve(data = brazil_route, 
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             # arrow = arrow(length = unit(0.02, "npc"), type = "closed"), 
             lty = 1, curvature = 0.3) + 
  labs(x = "Longitude",
       y = "Latitude") +
  theme_void()

```


### Graph comparing USA and USSR growth:

Dotted lines

```{r}
df_mad_usa_fussr <- df_mad %>% 
  filter(country_name %in% c("United States", "Former USSR"),
         indicator_code == "cgdppc",
         between(year, 1885, 1995))


df_mad_usa_fussr %>% 
  ggplot() +
  geom_line(aes(year, stat, lty = country_name), cex = 1) +
  geom_text(data = df_mad_usa_fussr %>%
              filter(country_name %in% c("United States", "Former USSR"),
              indicator_code == "cgdppc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.1)) +
  scale_y_continuous(labels = scales::comma_format()) + 
  expand_limits(x = 2010) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       # title = "GDP per capita growth since 1960",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "none")

```

Grey lines. I can easily cut the series before the dip in the 1990s - if that's preferable?

```{r}
# colour brewer greys

df_mad_usa_fussr %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name), cex = 1) +
  geom_text(data = df_mad_usa_fussr %>%
              filter(country_name %in% c("United States", "Former USSR"),
              indicator_code == "cgdppc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.1)) +
  scale_y_continuous(labels = scales::comma_format()) + 
  scale_color_grey(end = .5) +  
  expand_limits(x = 2010) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       title = "GDP per capita growth since 1960",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "none")

```

Dotted grey lines.

```{r}
df_mad_usa_fussr %>% 
  ggplot() +
  geom_line(aes(year, stat, colour = country_name, lty = country_name), cex = 1) +
  geom_text(data = df_mad_usa_fussr %>%
              filter(country_name %in% c("United States", "Former USSR"),
              indicator_code == "cgdppc",
         year == max(year)),
    aes(year, stat, label = country_name, hjust = -.1)) +
  scale_y_continuous(labels = scales::comma_format()) + 
  scale_color_grey(end = .5) +
  expand_limits(x = 2010) +
  labs(y = "GDP per capita (constant 2010 US$)", 
       x = "Year",
       title = "GDP per capita growth since 1960",
       caption = "Source: Maddison Project Database (2018)") +
  theme(legend.position = "none")

```



### Why does isiXhosa have clicks? The Bantu migration

```{r}
p_load(tidygeocoder, igraph)

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

arrow_areas <- c("yaounde, cameroon",
             "kigali, rwanda",
            "huambo, angola") %>% as_tibble()

arrow_areas_2 <- c("shinyanga, tanzania",
             "masasi, tanzania",
            "lilongwe, malawi",
            "cuvelai, angola") %>% as_tibble()

arrow_areas_3 <- c("mbeya, tanzania",
                    "Pietermaritzburg, south africa",
                    "pretoria, south africa",
                    "windhoek, namibia") %>% as_tibble()



# arrow_areas_coords <- arrow_areas %>% geocode(address = value, method = "osm", lat = lat1, long = long1)
# 
# write_rds(arrow_areas_coords, "data/arrow_areas_coords.rds")

arrow_areas_coords <- read_rds("data/arrow_areas_coords.rds")


# arrow_areas_2_coords <- arrow_areas_2 %>% geocode(address = value, method = "osm", lat = lat1, long = long1)
# 
# write_rds(arrow_areas_2_coords, "data/arrow_areas_2_coords.rds")

arrow_areas_2_coords <- read_rds("data/arrow_areas_2_coords.rds")


# arrow_areas_3_coords <- arrow_areas_3 %>% geocode(address = value, method = "osm", lat = lat2, long = long2)
# 
# write_rds(arrow_areas_3_coords, "data/arrow_areas_3_coords.rds")

arrow_areas_3_coords <- read_rds("data/arrow_areas_3_coords.rds")


arrow_areas_3_coords <- arrow_areas_3_coords %>% 
  rename(value1 = value)

arrow_directions_2 <- cbind(arrow_areas_2_coords, arrow_areas_3_coords) %>% 
  as_tibble() %>% select(-value, -value1)

yaonde <- arrow_areas_coords %>% filter(value == "yaounde, cameroon")

arrow_directions_1 <- arrow_areas_coords %>%
  filter(!value == "yaounde, cameroon") %>% 
  mutate(lat2 = yaonde$lat1,
         long2 = yaonde$long1) %>% 
  select(-value)

bm_labels <- tibble(
  x = c(22, 36, 12),
  y = c(-3, -25, 9),
  label = c("Spread of Bantu \n by 3000 years BP", "Spread of Bantu \n by 1500 years BP", "Bantu Homeland \n by 4000 years BP"))
```

Map of Bantu migration with lat and long lines

```{r}
ggplot(data = world) +
  geom_point(data = arrow_areas_coords, aes(long1, lat1)) +
  geom_sf(colour = "white") +
  coord_sf(ylim=c(-35,10), xlim = c(-10,50)) +
  geom_curve(data = arrow_directions_1, 
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             curvature = -.3,
             cex = 1.2,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 1, colour = "#a9a9a9") +
    geom_curve(data = arrow_directions_2, 
             aes(y = lat1, x = long1, yend = lat2, xend = long2),
             curvature = .3,
             cex = 1.2,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 2) +
  geom_point(data = yaonde, aes(long1, lat1), shape = 19, cex = 7) +
  geom_text_repel(data = bm_labels, aes(x, y, label = label), cex = 4) + 
  labs(x = "Longitude",
       y = "Latitude")

```

Without lat and long axes.
This is a simplified version of the map that only includes the arrows, and copying the text from the original.
Should this say "Bantu people" instead of "Bantu"?

```{r}
ggplot(data = world) +
  geom_point(data = arrow_areas_coords, aes(long1, lat1)) +
  geom_sf(colour = "white") +
  coord_sf(ylim=c(-35,10), xlim = c(-10,50)) +
  geom_curve(data = arrow_directions_1, 
             aes(y = lat2, x = long2, yend = lat1, xend = long1),
             curvature = -.3,
             cex = 1.2,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 1, colour = "#a9a9a9") +
    geom_curve(data = arrow_directions_2, 
             aes(y = lat1, x = long1, yend = lat2, xend = long2),
             curvature = .3,
             cex = 1.2,
             arrow = arrow(length = unit(0.02, "npc"), type = "closed"), lty = 2) +
  geom_point(data = yaonde, aes(long1, lat1), shape = 19, cex = 7) +
  geom_text_repel(data = bm_labels, aes(x, y, label = label), cex = 4) +
  theme_void()

```

